Synthetic algebraic geometry and the phenomenon of nongeometric sequents

### 2018-04-09 slides-pssl103.tex
Minor changes after the talk

  \vspace*{-2em}
  \hfill\video{zeros-in-families.mp4}{2cm}
\end{frame}

@@

  A \hil{synthetic affine scheme} is a set which is in bijection
  with~$\Spec(R)$ for some finitely presented algebra~$R$.

@@

\begin{frame}{Properties of the affine line}
  \begin{itemize}
    \item $\affl$ is a local ring:
    \[ 1 \neq 0 \qquad\qquad \text{$x + y$ \inv} \Longrightarrow \text{$x$ \inv} \vee
    \text{$y$ \inv} \]
    \item $\affl$ is a field:
    \begin{align*}
      \neg(x = 0) &\Longleftrightarrow \text{$x$ invertible} \\
      \neg(\text{$x$ invertible}) &\Longleftrightarrow \text{$x$ nilpotent}
    \end{align*}
    \item $\affl$ satisfies the axiom of microaffinity: Any map $f : \Delta \to
    \affl$ is of the form~$f(\varepsilon) = a + b \varepsilon$ for unique
    values~$a,b \? \affl$, where~$\Delta = \{ \varepsilon \? \affl \,|\,
    \varepsilon^2 = 0 \}$.
    \item Any function $\affl \to \affl$ is a polynomial.
    \item $\affl$ is anonymously algebraically closed: Any monic polynomial
    does \emph{not not} have a zero.
    \item $\affl$ is of unbounded Krull dimension.
  \end{itemize}
\end{frame}

@@

  Any known property of~$\affl$ follows from this \hil{synthetic
  quasicoherence}.

@@

  \begin{itemize}
    \item Additional \hil{nongeometric sequents} may hold in a classifying
    topos, for instance~``$\affl$ is synthetically quasicoherent''
    in~$\Zar(S)$. \\[-1em]
    \item These are \hil{$\boldsymbol{\TT}$-redundant}, but the converse is
    false. \\[-1em]  % (Bezem--Buchholtz--Coquand 2017). \\[-1em]
    \item \ \\[-1.2em]\mbox{Are they precisely the consequences of synthetic
    quasicoherence?} \\[-1em]
    \item Applications: synthetic algebraic geometry,
    generic freeness, \ldots
  \end{itemize}
\end{frame}

>>>>

  \vspace*{-2em}
  \hfill\video{zeros-in-families.mp4}{2cm} \\
  \hfill \scriptsize
  (see
  \textcolor{red}{\href{https://rawgit.com/iblech/internal-methods/master/images/zeros-in-families.mp4}{video}}
  for counterexample)
\end{frame}

@@

  A \hil{synthetic affine scheme} is a set which is in bijection
  with~$\Spec(R)$ for some synthetically quasicoherent~$\affl$-algebra~$R$.

@@

\begin{frame}{Properties of the affine line}
  \begin{itemize}
    \item $\affl$ is a local ring:
    \[ 1 \neq 0 \qquad\qquad \text{$x + y$ \inv} \Longrightarrow \text{$x$ \inv} \vee
    \text{$y$ \inv} \]
    \item $\affl$ is a field:
    \begin{align*}
      \neg(x = 0) &\Longleftrightarrow \text{$x$ invertible \quad [Kock 1976]} \\
      \neg(\text{$x$ invertible}) &\Longleftrightarrow \text{$x$ nilpotent}
    \end{align*}
    \item $\affl$ satisfies the axiom of microaffinity: Any map $f : \Delta \to
    \affl$ is of the form~$f(\varepsilon) = a + b \varepsilon$ for unique
    values~$a,b \? \affl$, where~$\Delta = \{ \varepsilon \? \affl \,|\,
    \varepsilon^2 = 0 \}$.
    \item Any function $\affl \to \affl$ is a polynomial.
    \item $\affl$ is anonymously algebraically closed: Any monic polynomial
    does \emph{not not} have a zero.
    \item $\affl$ is of unbounded Krull dimension.
  \end{itemize}
\end{frame}

@@

  Any known property of~$\affl$ follows from this \\
  \hil{synthetic quasicoherence}.

@@

  \begin{itemize}
    \item Additional \hil{nongeometric sequents} may hold in a classifying
    topos, for instance~``$\affl$ is synthetically quasicoherent''
    in~$\Zar(S)$. \\[-1em]
    \item These are \hil{$\boldsymbol{\TT}$-redundant}, but the converse is
    false. [Bezem--Buchholtz--Coquand 2017; answering a question by Wraith
    possibly raised at PSSL 1.] \\[-1em]
    \item \ \\[-1.2em]\mbox{Are they precisely the consequences of synthetic
    quasicoherence?} \\[-1em]
    \item Applications: synthetic algebraic geometry,
    generic freeness, \ldots
  \end{itemize}
\end{frame}
